Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-23 response-to-reviews.Rmd

I was eager to see the part of the manuscript about photoprotection and also its relationship to heat damage. The opening sentence (l. 384) was obvious to the point of being painful to read.  Of course photoprotection is higher in the upper canopy: radiation levels are higher! Can this first sentence get revised to increase information content? This is one of the more exciting topics of the review, yet is handled only in a cursory fashion. If the authors could manage a bit more discussion of this topic it would be useful – instead they move (in the same paragraph) to VOC emissions, something I view as a separate capability of some plants. This is also confused in Fig. 1 where VAZ and VOC are placed on the same line. I’d like to see a more complete discussion of photoprotection and heat dissipation by NPQ first, and a separate paragraph about VOC. Also, there is nothing about critical temperature until late in the manuscript even though the authors skirt the issue in their text about photoprotection. This kind of thermal sensitivity is different from gas exchange (the section where the text about Tcrit occurs, l. 530-555), and should have been discussed earlier in the manuscript.

**We seperated VAZ section from VOC, and provided further context to NPQ in the following lines:**

- **Efficient light-use is optimized through this process while preventing the accumulation of excess light, provoking the over production of harmful reactive oxygen species (ROS). However, at a photon flux density (PPFD) of $200\mu molm^{-2}s^{-1}$, within beech canopy, shaded lower canopy leaves have been observed to have higher NPQ rates than upper canopy reflecting rapid photosynthetic kinetics in shaded leaves as an acclimation to sunflecks. However, beyond this PPFD threshold, NPQ rates of upper canopy increase with an increase in light intensity [@scartazzaInvestigatingEuropeanBeech2016a]. The dissipation of excess light energy is also important for leaf acclimation to higher temperatures if high $T_{leaf}$ impairs the photochemical energy dissipation pathway [@havauxTemperaturedependentAdjustmentThermal1996].**

### 2022-01-23 vertical-thermal-review.Rmd
seperating paragraphs: biochemical and VOC
Added NPQ lines to photoprotection and responded to comment on NPQ

Biochemical protection against foliage light and heat damage also increases with irradiance, and thus tends to be higher in the upper canopy than in the understory (Table `r table_leaf_traits`). 
Carotenoids (e.g., beta carotene and lutein), including xanthophyll cycle pigments (i.e., violaxanthin, antheraxanthin and zeaxanthin, VAZ), can play a role in antioxidant scavenging and converting excess excitation energy into heat to reduce photoinhibition [@niinemetsPhotosynthesisResourceDistribution2007; @mathurPhotosyntheticEfficiencySun2018a; @niinemetsAcclimationHighIrradiance1998]. 
The dissipation of excess light energy is also important for leaf acclimation to higher temperatures if high $T_{leaf}$ impairs the photochemical energy dissipation pathway [@havauxTemperaturedependentAdjustmentThermal1996]. 
Capacity for heat-sensitive, light-dependent (photosynthetically linked) emissions of volatile organic compounds (VOCs), including isoprene and monoterpenes, enhances photosynthetic thermal tolerance by regulating antioxidant defenses and other metabolic processes [@sharkeyIsopreneEmissionPlants2008; @taylorCapacityEmitIsoprene2019; @monsonLeafIsopreneEmission2021; @vickersUnifiedMechanismAction2009; @copoloviciCapacityThermalProtection2005; @riedlmeierMonoterpenesSupportSystemic2017]. 
Less than half of tree species express significant light-dependent emissions [@kesselmeierBiogenicVolatileOrganic; @taylorIsopreneEmissionStructures2018], such that variation in emissions across the vertical profile can be influenced both by vertical variation in species compositions and plasticity in emission rate capacities [see section 4.4, @taylorNewFieldInstrument2021].

>>>>

Biochemical protection against foliage light and heat damage also increases with irradiance, and thus tends to be higher in the upper canopy than in the understory (Table `r table_leaf_traits`). 
Carotenoids (e.g., beta carotene and lutein), including xanthophyll cycle pigments (i.e., violaxanthin, antheraxanthin and zeaxanthin, VAZ), can play a role in antioxidant scavenging and converting excess excitation energy into heat to reduce photoinhibition, and therefore also serve as thermoprotection and postpone heat damage [@niinemetsPhotosynthesisResourceDistribution2007; @mathurPhotosyntheticEfficiencySun2018a; @niinemetsAcclimationHighIrradiance1998]. With saturating irradiance in canopy leaves, unused energy from photosystem II antennae is further dissipated as heat energy through non-photochemical quenching (NPQ) [@mathurPhotosyntheticEfficiencySun2018b]. Efficient light-use is optimized through this process while preventing the accumulation of excess light, provoking the over production of harmful reactive oxygen species (ROS). However, at a photon flux density (PPFD) of $200\mu molm^{-2}s^{-1}$, within beech canopy, shaded lower canopy leaves have been observed to have higher NPQ rates than upper canopy reflecting rapid photosynthetic kinetics in shaded leaves as an acclimation to sunflecks. However, beyond this PPFD threshold, NPQ rates of upper canopy increase with an increase in light intensity [@scartazzaInvestigatingEuropeanBeech2016a]. The dissipation of excess light energy is also important for leaf acclimation to higher temperatures if high $T_{leaf}$ impairs the photochemical energy dissipation pathway [@havauxTemperaturedependentAdjustmentThermal1996].

Capacity for heat-sensitive, light-dependent (photosynthetically linked) emissions of volatile organic compounds (VOCs), including isoprene and monoterpenes, enhances photosynthetic thermal tolerance by regulating antioxidant defenses and other metabolic processes [@sharkeyIsopreneEmissionPlants2008; @taylorCapacityEmitIsoprene2019; @monsonLeafIsopreneEmission2021; @vickersUnifiedMechanismAction2009; @copoloviciCapacityThermalProtection2005; @riedlmeierMonoterpenesSupportSystemic2017]. 
Less than half of tree species express significant light-dependent emissions [@kesselmeierBiogenicVolatileOrganic; @taylorIsopreneEmissionStructures2018], such that variation in emissions across the vertical profile can be influenced both by vertical variation in species compositions and plasticity in emission rate capacities [see section 4.4, @taylorNewFieldInstrument2021].
